---
title: "Analyses - MPA by country"
author : "Nina Prasil & Jeremy Wicquart"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: "cosmo"
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, fig.align = "center")

```


## A DESTINATION DU PRINCE ##
```{r}

##### 1. Required packages ----

library(tidyverse) # Core tidyverse packages
library(readxl) # To read excel files
library(sf) # To plot maps
sf_use_s2(FALSE) # Switch from S2 to GEOS

##### 2. Load data ----

#2.1. Reef polygons data

load(file = "./../data/reefs.eez.RData")


data_reefs <- reefs.eez %>%
  # Re-assign TERRITORY1 depending on geopolitical issues
  mutate(SOVEREIGN1 = case_when(TERRITORY1 == "Glorioso Islands" ~ "France",
                                TERRITORY1 == "Chagos Archipelago" ~ "United Kingdom",
                                TRUE ~ SOVEREIGN1)) %>% 
  st_transform(., crs = 32662)


# 2.2 Reef surface data

reef_total <- read_csv2("../figs/04_table-reef-surface-by-country.csv")

# 2.3 MPA data
# Total surface of MPA per country
mpa_tot_area <- read_csv2("../figs/02_mpa-with-reef.csv")

# Unioned MPA polygons
load("./../data/02_mpa/00-data_mpa_union.RData")


## 3. Overlap reefs and MPA
# Intersection coral reefs and MPA

protected_reefs <- data_reefs %>% 
  filter(SOVEREIGN1 %in% top_30) %>% 
  arrange(SOVEREIGN1) %>% 
  st_intersection(., total_mpa) %>% 
  mutate(cr_mpa = st_area(geometry)) %>% 
  st_drop_geometry() %>% 
  group_by(SOVEREIGN1) %>% 
  summarise(cr_in_mpa = sum(cr_mpa))

units(protected_reefs$cr_in_mpa) <- "km2"  

# Percentage of coral reefs under protection

protected_reefs <- protected_reefs %>% 
  left_join(., reef_total) %>% 
  group_by(SOVEREIGN1) %>% 
  mutate(cr_in_mpa_rel = cr_in_mpa/total_area_abs*100)

# Export cr_in_mpa 
write_csv2(protected_reefs, "./../data/02_mpa/00_protected-coral-reefs.csv")


```




# Import functions, packages and data
- MPA zones implemented in the 30 countries studied
- Data filtration : * Were retained only protected areas fitting the "is MAP" definition (as defined by MPAtlas) 
* Are taken into account only MPA zones that are considered implemented (as defined by MPAtlas)

```{r base}

source("functions/graphical_par.R")
source("functions/theme_map.R")


##### 2. Required packages ----

library(tidyverse) # Core tidyverse packages
library(readxl) # To read excel files
library(sf) # To plot maps
sf_use_s2(FALSE) # Switch from S2 to GEOS

##### 3. Load data ----

### 3.1? EEZ from Falnders Institute 
data_bbox <- read_xlsx("./../data/top_30_bbox.xlsx", sheet = 1)

data_eez <- read_sf("../data/01_maps/World_EEZ_v11_20191118/eez_v11.shp") %>% 
  # Re-assign Territory depending to geopolitical issues
  mutate(SOVEREIGN1 = case_when(TERRITORY1 == "Glorioso Islands" ~ "France",
                                TERRITORY1 == "Chagos Archipelago" ~ "United Kingdom",
                                TRUE ~ SOVEREIGN1)) %>% 
  # Extract top 30 countries
  filter(SOVEREIGN1 %in% data_bbox$SOVEREIGN1) %>% 
  # Remove territories that do not contains coral reefs 
  filter(!(TERRITORY1 %in% c("Kerguélen", "Falkland / Malvinas Islands", "United Kingdom", "France",
                             "Saint Helena", "Alaska", "Crozet Islands", "Amsterdam and Saint Paul Islands",
                             "South Georgia and the South Sandwich Islands", "Jersey", "Guernsey", "French Guiana",
                             "Heard and McDonald Islands", "Macquarie Island", "Ascension",
                             "Saint-Pierre and Miquelon", "Tristan da Cunha", "Gibraltar"))) %>% 
  st_transform(., crs = 32662)


### 3.2. MPAs from MPAtlas

data_mpa <- read_sf("./../data/02_mpa/mpatlas_20201223_clean/mpatlas_20201223_clean.shp") %>% 
  # Select top 30
  filter(sovereign %in% c("AUS", "IDN", "PHL", "FRA", "PNG", "USA", "FJI", "SLB", "SAU", "MDV",
                          "CUB", "FSM", "BHS", "MDG", "MHL", "IND", "MYS", "EGY", "KIR", "TZA",
                          "MOZ", "SDN", "ERI", "JPN", "SYC", "BLZ", "GBR", "VUT", "TON", "MEX")) %>% 
  # Homogenize countries names
  rename(SOVEREIGN1 = sovereign) %>% 
  filter(is_mpa == 1) %>% 
  filter(implemente == 1) %>% 
  mutate(SOVEREIGN1 = str_replace_all(SOVEREIGN1, 
                                      c("AUS" = "Australia",
                                        "BHS" = "Bahamas", 
                                        "BLZ" = "Belize",
                                        "CUB" = "Cuba", 
                                        "EGY" = "Egypt", 
                                        "ERI" = "Eritrea", 
                                        "FJI" = "Fiji", 
                                        "FRA" = "France", 
                                        "FSM" = "Micronesia",
                                        "GBR" = "United Kingdom",
                                        "IDN" = "Indonesia", 
                                        "IND" = "India", 
                                        "JPN" = "Japan", 
                                        "KIR" = "Kiribati",
                                        "MDG" = "Madagascar", 
                                        "MDV" = "Maldives", 
                                        "MEX" = "Mexico",
                                        "MHL" = "Marshall Islands",
                                        "MOZ" = "Mozambique", 
                                        "MYS" = "Malaysia",
                                        "PHL" = "Philippines", 
                                        "PNG" = "Papua New Guinea",
                                        "SAU" = "Saudi Arabia",
                                        "SDN" = "Sudan", 
                                        "SLB" = "Solomon Islands",
                                        "SYC" = "Seychelles",
                                        "TON" = "Tonga",
                                        "TZA" = "Tanzania",
                                        "USA" = "United States",
                                        "VUT" = "Vanuatu"))) %>% 
  st_transform(., crs = 32662) %>%
  st_simplify(., dTolerance = 50) %>% 
  st_intersection(., data_eez) 
  

data_mpa <- data_mpa %>% st_make_valid(.)


### 3.3. Transformed data from WRI - Reefs merged with EEZ

load(file = "./../data/reefs.eez.RData")


data_reefs <- reefs.eez %>%
  # Re-assign TERRITORY1 depending on geopolitical issues
  mutate(SOVEREIGN1 = case_when(TERRITORY1 == "Glorioso Islands" ~ "France",
                                TERRITORY1 == "Chagos Archipelago" ~ "United Kingdom",
                                TRUE ~ SOVEREIGN1)) %>% 
  st_transform(., crs = 32662)



```

# Surface MPA

```{r}

# 1. Calculate MPA area

mpa_area <- list()

for (i in unique(data_mpa$SOVEREIGN1)) {
  
  data_mpa_i <- data_mpa %>% 
    filter(SOVEREIGN1 == i) 
  
  mpa_comb_i <- st_union(st_combine(data_mpa_i), by_feature = T)
   
  mpa_area_i <- st_area(mpa_comb_i) 
  
  units(mpa_area_i) <- "km^2"
  
  mpa_area[[i]] = mpa_area_i
  
}



# 2. Clean table with all MPA related information

mpa_table <- as.data.frame(mpa_area, row.names = T) %>% 
  
  pivot_longer(., "United.States":"Indonesia", names_to = "SOVEREIGN1", values_to = "mpa_area_km2") %>% 
  mutate(SOVEREIGN1 = str_replace_all(SOVEREIGN1, 
                                      c("Marshall.Islands" = "Marshall Islands",
                                        "Papua.New.Guinea" = "Papua New Guinea",
                                        "Saudia.Arabia" = "Saudi Arabia",
                                        "Solomon.Islands" = "Solomon Islands",
                                        "United.Kingdom" = "United Kingdom",
                                        "United.States" = "United States"))) %>% 
  arrange(SOVEREIGN1) %>% 
  mutate(mpa_area_km2 = round(mpa_area_km2, 0 ))

# 3. Save table 

write_csv2(mpa_table, "../figs/02_mpa-with-reef.csv")


mpa_tot_area <- read_csv2("../figs/02_mpa-with-reef.csv")

```

# 2. Reef surface under protection
Tjrs le meme pb avec l'Australie GRRRRRRRRRRRRRRRRRRRRRR
-> Australia removed and it works

```{r}


#### 1. Dissolve all MPA zones in one polygon per country

#top_30 <- c("Belize", "Egypt")

top_30 <- unique(data_mpa$SOVEREIGN1)

  
# Loop to create the combined polygons for each country
mpa_tot <- list()

for (i in top_30) {
  
  data_mpa_i <- data_mpa %>% 
    filter(SOVEREIGN1 == i) 
    
  mpa_poly_i <- st_union(st_combine(data_mpa_i), by_feature = TRUE) 
  
  mpa_tot[[i]] <- mpa_poly_i
  
}

# Convert the list in a table
total_mpa <- as_tibble(mpa_tot[1:29], rownames = NULL) %>% 
  # From the 1st to the last country in top_30
  pivot_longer(., "United States": "Indonesia", names_to = "SOVEREIGN1", values_to = "geometry") %>% 
  st_as_sf(., crs = 32662) %>% 
  st_transform(., crs = 32662) %>% 
  arrange(SOVEREIGN1)

# Export mpa_tot, since spatial analyses are time consuming
save(mpa_tot, 
     file = "./../data/02_mpa/00-data_mpa_merged_2.RData")

save(total_mpa, 
     file = "./../data/02_mpa/00-data_mpa_union.RData")

load("./../data/02_mpa/00-data_mpa_union.RData")

#### 2. Coral reefs intersection with MPA

cr_in_mpa <- data_reefs %>% 
  filter(SOVEREIGN1 %in% top_30) %>% 
  arrange(SOVEREIGN1) %>% 
  st_intersection(., total_mpa) %>% 
  mutate(cr_mpa = st_area(geometry)) %>% 
  st_drop_geometry() %>% 
  group_by(SOVEREIGN1) %>% 
  summarise(mpa_cr_ctry = sum(cr_mpa))

units(cr_in_mpa$mpa_cr_ctry) <- "km2"  

# Export cr_in_mpa, since spatial analyses are time consuming
write_csv2(cr_in_mpa, "./../data/02_mpa/00_protected-coral-reefs.csv")


test <- cr_in_mpa %>% 
  ungroup() %>% 
  summarise(total = sum(mpa_cr_ctry))
```


# Australia

```{r}

# MPA Australia
mpa_aus <- data_mpa %>% 
  filter(SOVEREIGN1 == "Australia") 

test <- st_simplify(mpa_aus, dTolerance = 100)


union_mpa <- st_union(st_combine(test), by_feature = TRUE)

temp <- total_mpa %>% filter(SOVEREIGN1 == "Egypt") 
temp_bis <- data_reefs %>% filter(SOVEREIGN1 == "Egypt")

ggplot() +
  geom_sf(data = temp_bis) +
  geom_sf(data = temp_bis, col = "blue") +
  geom_sf(data = test, col = "red")

mpa_aus_km2 <- st_area(union_mpa)
units(mpa_aus_km2) <- "km2"

# Intersection coral reefs and MPA

cr_mpa_aus <- data_reefs %>% 
  filter(SOVEREIGN1 == "Australia") %>% 
  st_intersection(., union_mpa) %>% 
  mutate(cr_mpa = st_area(geometry)) %>% 
  st_drop_geometry() %>% 
  group_by(SOVEREIGN1) %>% 
  summarise(mpa_cr_ctry = sum(cr_mpa))

# Units from m2 to km2
units(cr_mpa_aus$mpa_cr_ctry) <- "km2"


```



# Not used - just in case - Coral reefs overlap 

```{r}

# Select top 30 countries (29, since Eritrea does not have MPA)
top_30 <- unique(data_mpa$SOVEREIGN1)

# Select the coral reefs in MPA and calculate their surface per country
mpa_w_cr <- data_reefs %>% 
  filter(SOVEREIGN1 %in% top_30) %>% 
  # create polygons of overlaps btw coral reefs and MPA zones
  st_intersection(., st_make_valid(data_mpa)) %>% 
  # Calculate the area of reefs under protection
  group_by(SOVEREIGN1) %>% 
  mutate(cr_mpa = st_area(geometry)) %>% 
  #summarise(mpa_cr_ctry = sum(cr_mpa)) %>% 
  #hopefully getting rid of the overlaps
  distinct(Reef_id, .keep_all = TRUE) %>% 
  summarise(mpa_cr_ctry = sum(cr_mpa))

# Convert m^2 in km^2
units(mpa_w_cr$mpa_cr_ctry) <- "km^2"

# 3. Export to RData

save(mpa_w_cr, 
     file = "./../data/02_mpa/00-mpa_cr_area_merged.RData")

############## TESTS #############
# pb: Australia: reef area in MPA > total reef area
temp <- data_reefs %>% 
  filter(SOVEREIGN1 == "Indonesia") %>% 
  st_intersection(., st_make_valid(data_mpa))  %>% 
  mutate(cr_mpa = st_area(geometry)) %>%
  st_drop_geometry() %>% 
  distinct(Reef_id, .keep_all = TRUE) %>% 
  summarise(mpa_cr_ctry = sum(cr_mpa)/1000000)


temp <- mpa_w_cr %>% 
  filter(SOVEREIGN1 == "Australia")


  ggplot() +
  geom_sf(data = test, col = "blue") +
    geom_sf(data = test2, col = "green") +
  geom_sf(data = temp, col = "red") 
  

  

#######
test <- data_mpa %>% 
  filter(SOVEREIGN1 %in% c("Australia"))
  
test2 <- data_reefs %>% 
  filter(SOVEREIGN1 == "Australia")
  
temp <- data_reefs %>% 
  filter(SOVEREIGN1 == "Australia") %>% 
  mutate(cr_tot = st_area(geometry)) %>% 
  summarise(cr_ctry = sum(cr_tot))


#######
ctry_test <- c("Belize", "Egypt", "Sudan")

for (i in ctry_test) {
  

  
}

  test <- st_join(data_reefs, temp, left = FALSE) %>% 
  st_drop_geometry() %>% 
  distinct(., mpa_id)



ggplot() +
  #geom_sf(data = temp) +
  geom_sf(data = mpa_w_cr)

```




# Reproducibility

```{r reprod}

# 1. Reproducibility ----

sessionInfo()

```

---
Nina PRASIL & Jeremy WICQUART | `r format(Sys.time())`